### 2014-11-12 related.tex
Update related work.

\paragraph{Declarative Languages}
Declarative programming paradigms such as \emph{logic} and \emph{functional}
programming have been extensively exploited for their implicit parallelism.
In logic languages such as Prolog, researchers took advantage of the
non-determinism of proof-search to evaluate subgoals in parallel with models
such as \emph{or-parallelism}~\cite{ali-86} and
\emph{and-parallelism}~\cite{Shen-92}.  In functional languages, the stateless
nature of computation allows multiple expressions to evaluate safely in
parallel. This has been explored in several languages such as
NESL~\cite{Blelloch:1996:PPA:227234.227246}, Id~\cite{Nikhil93anoverview}, and
more recently Data Parallel Haskell~\cite{nepal}.

\paragraph{Data Centric Languages}
Recently, there has been an increasing interest in declarative data-centric
languages.  MapReduce~\cite{Dean:2008:MSD:1327452.1327492}, for instance, is a
popular data-centric programming model that is optimized for large clusters.
Intrinsic to its popularity is the simplicity of its scheduling and data sharing
model.  In order to facilitate the writing of programs over large datasets,
SQL-like languages such as PigLatin~\cite{Olston:2008:PLN:1376616.1376726} have
been developed.  PigLatin builds on top of MapReduce and allows the programmer
to write complex data-flow graphs, raising the abstraction and easy of
programmability of MapReduce programs.  An alternative to PigLatin/MapReduce is
Dryad~\cite{Isard:2007:DDD:1272996.1273005} that allows programmers to design
arbitrary computation patterns using the DAG abstraction. It combines
computational vertices with communication channels (edges) that are
automatically scheduled to run on multiple computers/cores.

GraphLab~\cite{GraphLab2010} is a C++ framework for developing parallel machine
learning algorithms. GraphLab allows nodes to have read/write access to
different scopes through different concurrent access models in order to balance
performance and data consistency.  GraphLab also provides different schedulers
that dictate the order in which node's are computed, allowing the programmer to
optimize the program. Later in this paper, we will show how one GraphLab
scheduler can be implemented in CLM through the use of coordination facts.

>>>>

\paragraph{Declarative Languages}
Declarative programming paradigms such as \emph{logic} and \emph{functional}
programming have been extensively exploited for their implicit parallelism.
In logic languages such as Prolog, researchers took advantage of the
non-determinism of proof-search to evaluate subgoals in parallel with models
such as \emph{or-parallelism}~\cite{ali-86} and
\emph{and-parallelism}~\cite{Shen-92}.  In functional languages, the stateless
nature of computation allows multiple expressions to evaluate safely in
parallel. This has been explored in several languages such as
NESL~\cite{Blelloch:1996:PPA:227234.227246}, Id~\cite{Nikhil93anoverview}, and
more recently Data Parallel Haskell~\cite{nepal}.
NESL often got good performance, but was limited to a particular
application space. We also limit our application space, but also give the
programmer a declarative way to coordinate computation to improve the
performance of the program.

\paragraph{Data Centric Languages}
Recently, there has been an increasing interest in declarative data-centric
languages. MapReduce~\cite{Dean:2008:MSD:1327452.1327492}, for instance, is a
popular programming model that is optimized for large clusters.
Intrinsic to its popularity is the simplicity of its scheduling and data sharing
model.  In order to facilitate the writing of programs over large datasets,
SQL-like languages such as PigLatin~\cite{Olston:2008:PLN:1376616.1376726} have
been developed.  PigLatin builds on top of MapReduce and allows the programmer
to write complex data-flow graphs, raising the abstraction and easy of
programmability of MapReduce programs.  An alternative to PigLatin/MapReduce is
Dryad~\cite{Isard:2007:DDD:1272996.1273005} that allows programmers to design
arbitrary computation patterns using the DAG abstraction. It combines
computational vertices with communication channels (edges) that are
automatically scheduled to run on multiple computers/cores.

GraphLab~\cite{GraphLab2010} is a C++ framework for developing graph-based parallel machine
learning algorithms. GraphLab allows nodes to have read/write access to
different scopes through different concurrent access models in order to balance
performance and data consistency.  GraphLab also provides different schedulers
that dictate the order in which node's are computed, allowing the programmer to
optimize the program. Later in this paper, we will show how one GraphLab
scheduler can be implemented in CLM through the use of coordination facts.
Another, more restrictive, graph-based system is
Pregel~\cite{Malewicz:2010:PSL:1807167.1807184}, where graph algorithms must be
executed as a sequence of iterations of computation and message passing.
Both those systems provide excellent performance, however they are not
declarative and have a steep learning curve.
